---
title: 'Sensing Invisible Others: The Case of Leland Stanford Junior'
author: "Eleanor Schille-Hudson" 
date: "2024-10-23"
output:
  html_document:
    toc: yes
    df_print: paged
    toc_float: yes
  html_notebook:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

```{r global_options, include = F}
#setting up file
knitr::opts_chunk$set(echo=F, warning=F, cache=F, message=F)
```

```{r}
# loading necessary packages
library(tidyverse)
library(readxl)
library(ggpubr)
library(rstatix)
library(lme4)
library(lmerTest)
library(parameters)
library(ggsignif)
library(RColorBrewer)
```

These are the analyses which accompany the paper Sensing invisible others: the case of Leland Stanford Junior. Running this RMarkdown file will produce the results and plots reported in the manuscript. 

# Study 1 

```{r}
# loading data study 1
s1 <- read_excel("leland_study1.xlsx", sheet = "study1_cleaned") %>%
   mutate(Age = as.numeric(Age))
```

## Cleaning & Organizing

```{r}
# creating key 
key_s1 <- s1 %>% slice(1) %>% 
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "qualtrics_question", values_to = "label")

# removing that first row from s1
s1 <- s1[-1,]

# filtering out people who didn't answer one (or more) of the sets of questions
# this leaves us with 30 participants
s1 <- s1 %>% subset(!(IndivID %in% c(1103, 1257, 1699, 1785))) 
```

```{r}
# making absorption and Leland scores 
s1 <- s1 %>% 
  mutate(across((S1_Q10:S3_Q5_12), as.numeric)) %>% # making everything numeric again, many variables character type from labels
  mutate(absorption_score = rowSums(across(S1_Q10:S1_Q43), na.rm = T), # calculating absorption scores by summing
         leland_lecture_score = rowSums(across(c(S2_Q3_1:S2_Q3_5, S2_Q5_1:S2_Q5_8, S2_Q5_10, S2_Q5_12)), na.rm = T), # summing Leland lecture qs
         leland_tracks_score = rowSums(across(c(S3_Q3_1:S3_Q3_5, S3_Q5_1:S3_Q5_8, S2_Q5_10, S2_Q5_12)), na.rm = T)) #summing Leland tracks qs
```

```{r}
# making a dataframe where all responses are scaled from 0-1
s1_scaled <- s1 %>% mutate(across(c(S2_Q3_1:S2_Q3_5, S2_Q5_1:S2_Q5_8, S2_Q5_10, S2_Q5_12), ~./7),
                           S2_Q4_1 = (S2_Q4_1 - 1)/9,
                           across(c(S3_Q3_1:S3_Q3_5, S3_Q5_1:S3_Q5_8, S2_Q5_10, S2_Q5_12), ~./7),
                           S3_Q4_1 = (S3_Q4_1 - 1)/9)

# summing scaled Leland scores
s1_scaled <- s1_scaled %>%
  mutate(leland_lecture_score_scaled = rowSums(across(c(S2_Q3_1:S2_Q3_5, S2_Q5_1:S2_Q5_8, S2_Q5_10, S2_Q5_12)), na.rm = T), # summing Leland lecture qs
         leland_tracks_score_scaled = rowSums(across(c(S3_Q3_1:S3_Q3_5, S3_Q5_1:S3_Q5_8, S2_Q5_10, S2_Q5_12)), na.rm = T)) #summing Leland tracks qs
```

## T-Test

```{r}
# paired t-test on Leland questions MINUS peaceful question, but SCALED 
# maximum score of 15 (1pt/question, 15 questions)

## getting right data frame
s1_lelandt_scaled <- s1_scaled %>% 
  select(IndivID, leland_lecture_score_scaled, leland_tracks_score_scaled) %>% 
  pivot_longer(cols = -IndivID, names_to = "group", values_to = "score")

## summary stats
s1_lelandt_scaled %>%
    group_by(group) %>%
    get_summary_stats(score, type = "mean_sd")

## paired t-test
t_test(score ~ group, data = s1_lelandt_scaled, paired = TRUE)

## effect size
s1_lelandt_scaled  %>% cohens_d(score ~ group, paired = TRUE)
```

## Box Plot

```{r}
# box plot with lines
s1_lelandt_scaled %>%
  mutate(group = factor(group, labels = c("lecture condition", "tracks condition"))) %>% 
  ggplot( aes(x=group, y=score, fill=group)) +
    geom_boxplot() +
    geom_jitter(height = 0, width = 0.00,
      color="black", size=0.4, alpha=0.5) +
  geom_line(aes(group = IndivID), color = "grey", alpha = 0.5) +
  geom_signif(comparisons = list(c("lecture condition", "tracks condition")),
              annotations="***") +
    theme_bw() +
  scale_fill_manual(values = c("lecture condition" = "#00BFC4", "tracks condition" = "#F8766D")) +  # Swap lecture and tracks colors
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("Scaled Leland Scores for Lectures vs Tracks Conditions") +
    xlab("") +
  ylab("Scaled Leland Score")

```

## Correlation of Leland Scores and Absorption

```{r}
#correlation between absorption and lecture condition Leland score
lecture_abs_r <- cor.test(s1_scaled$absorption_score, s1_scaled$leland_lecture_score_scaled,
                             method = "pearson")
lecture_abs_r

#correlation between absorption and tracks condition Leland score
tracks_abs_r <- cor.test(s1_scaled$absorption_score, s1_scaled$leland_tracks_score_scaled,
                             method = "pearson")
tracks_abs_r
```

## Scatterplot

```{r, fig.width = 8}
# minus peace SCALED

s1_scaled %>% 
  select(IndivID, leland_lecture_score_scaled, leland_tracks_score_scaled, absorption_score) %>% 
  rename(`Lecture Score` = leland_lecture_score_scaled, `Tracks Score` = leland_tracks_score_scaled) %>%
  pivot_longer(cols = c(`Lecture Score`, `Tracks Score`), names_to = "group", values_to = "score") %>%
  ggplot(aes(x=absorption_score, y=score)) + 
  geom_jitter(
    aes(color = group),
    height = 0, width = 0, 
              size = 1, alpha = 1) +
  #geom_smooth(aes(group = NULL), color = "black", method = "lm") +
  stat_cor(method = "pearson") +
  facet_wrap(~ group, ncol = 2) + 
  scale_color_brewer(palette = "Set2") +
  labs(title = "Relationship between Scaled Leland Score and Absorption Score",
       subtitle = "Both Scores Summed",
       x = "Absorption Score",
       y = "Leland Score",
       color = "Group") +
  theme(legend.position="none") +
   guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)))
```

## Mixed Effect Models

```{r}
# model where score is predicted by group and absorption (and participant)

## data frame 
s1_m_np <- s1_scaled %>% 
  select(IndivID, leland_lecture_score_scaled, leland_tracks_score_scaled, absorption_score) %>% 
  pivot_longer(cols = c(leland_lecture_score_scaled, leland_tracks_score_scaled), names_to = "group", values_to = "score") %>%
  mutate(IndivID = factor(IndivID))

## mixed effects model without interaction 
s1.m <- lmer(score ~ scale(absorption_score, scale=F) + group + (1 | IndivID),
             data = s1_m_np)
parameters(s1.m)

## mixed effects model with interaction 
s1.m_int <- lmer(score ~ scale(absorption_score, scale=F) * group + (1 | IndivID),
             data = s1_m_np)
parameters(s1.m_int)

```

## Scatterplot with Regression 

```{r, fig.width = 8}

s1_scaled %>% 
  select(IndivID, leland_lecture_score_scaled, leland_tracks_score_scaled, absorption_score) %>% 
  rename(`Lecture Score` = leland_lecture_score_scaled, `Tracks Score` = leland_tracks_score_scaled) %>%
  pivot_longer(cols = c(`Lecture Score`, `Tracks Score`), names_to = "group", values_to = "score") %>%
  ggplot(aes(x=absorption_score, y=score)) + 
  geom_jitter(
    aes(color = group),
    height = 0, width = 0, 
              size = 1, alpha = 1) +
  geom_smooth(aes(color = group), method = "lm") +
  scale_color_brewer(palette = "Set2") +
  labs(title = "Relationship between Scaled Leland Score and Absorption Score",
       subtitle = "Both Scores Summed",
       x = "Absorption Score",
       y = "Leland Score",
       color = "Group") +
   guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)))
```

# Study 2

## Cleaning and Organizing

```{r}
# loading data study 2
s2_clean <- read_excel("leland_study2.xlsx", sheet = "study2_cleaned")
```

```{r}
# making data set for reported Leland score in Study 2
s2 <- s2_clean %>% 
  select(IndivID, group, gender, tellegentotal, 
         Ipod, playcount,
         pq1, pq1d, pq2, pq2a, pq3, pq3c, pq4, pq4b, pq6, pq6c, pq7, pq7b,
         pq11, pq21, pq22a, pq22b, pq22c, pq22d) %>% 
  filter(Ipod < 3) #removing Ipod 3 person, that was a mistake

# renaming, cleaning up
s2 <- s2 %>%
  mutate(across(pq1:pq22d, na_if, 99)) %>% #making all 99 values (means "missing" into NAs)
  rename(condition = Ipod) %>% 
  mutate(condition = factor(ifelse(condition == 1, "hearing", "seeing")), #making the conditions clearer
         leland_score = rowSums(across(pq1:pq22d), na.rm = T)) %>% #creating Leland score
  filter(playcount >= 2) %>% #filtering out people who listened less than 2x
  filter(leland_score > 0 | if_any(pq1:pq22d, ~ !is.na(.))) # exclude rows with all NAs in Leland columns
```

```{r}
# making dataset for Leland score that includes all questions (results reported in Appendix B)
s2_allq <- s2_clean %>% 
  select(IndivID, group, gender, tellegentotal, 
         Ipod, playcount,
         pq1, pq2, pq3, pq4, pq5, pq6, pq7, pq8, pq9, pq10,
         pq11, pq12, pq13, pq14, pq15, pq16, pq17, pq18, pq19, 
         pq20, pq21, pq22a, pq22b, pq22c, pq22d) %>% 
  filter(Ipod < 3) #removing Ipod 3 person, that was a mistake

# renaming, cleaning up
s2_allq <- s2_allq %>%
  mutate(across(pq1:pq22d, na_if, 99)) %>% #making all 99 values (means "missing" into NAs)
  rename(condition = Ipod) %>% 
  mutate(condition = factor(ifelse(condition == 1, "hearing", "seeing")), #making the conditions clearer
         leland_score = rowSums(across(pq1:pq22d), na.rm = T)) %>% #creating Leland score
  filter(playcount >= 2) %>% #filtering out people who listened less than 2x
  filter(leland_score > 0 | if_any(pq1:pq22d, ~ !is.na(.))) # exclude rows with all NAs in Leland columns
```

```{r}
# making dataset for Leland score that includes all questions and all follow-ups (results reported in Appendix B)
s2_allq_allf <- s2_clean %>% 
  select(IndivID, group, gender, tellegentotal, 
         Ipod, playcount,
         pq1, pq1b, pq1c, pq1d, 
         pq2, pq2a, 
         pq3, pq3b, pq3c, 
         pq4, pq4a, pq4b, 
         pq5, pq5a, pq5b, 
         pq6, pq6b, pq6c, 
         pq7, pq7a, pq7b, 
         pq8, pq9, pq10,
         pq11, pq11a, pq11b, 
         pq12, pq13, pq14, pq15, pq16, pq17, pq18, pq19, 
         pq20, pq21, pq22a, pq22b, pq22c, pq22d) %>% 
  filter(Ipod < 3) #removing Ipod 3 person, that was a mistake

# renaming, cleaning up
s2_allq_allf <- s2_allq_allf %>%
  mutate(across(pq1:pq22d, na_if, 99)) %>% #making all 99 values (means "missing" into NAs)
  rename(condition = Ipod) %>% 
  mutate(condition = factor(ifelse(condition == 1, "hearing", "seeing")), #making the conditions clearer
         leland_score_plus = rowSums(across(pq1:pq22d), na.rm = T)) %>% #creating Leland score
  filter(playcount >= 2) %>% #filtering out people who listened less than 2x
  filter(leland_score_plus > 0 | if_any(pq1:pq22d, ~ !is.na(.))) # exclude rows with all NAs in Leland columns
```

*Note* Here we are using the variable "tellegentotal." The summed scores in that column come from pen and paper survey responses, not all of which were broken down and put question-by-question into the data file. This means that we cannot fully reconstruct people's absorption responses on the question level, but that is ok. We trust the pen and paper totals. 

## T-Tests

### Leland score 

```{r}
# hearing vs. seeing--t-test reported in results of Study 2

## summary stats
s2 %>%
    group_by(condition) %>%
    get_summary_stats(leland_score, type = "mean_sd")

## survey questions 
t_test(leland_score ~ condition, data = s2)

## absorption
t_test(tellegentotal ~ condition, data = s2)
```


### Special seeing and hearing Leland subscores

```{r}
# did seeing people see more leland, hear more leland?

## dataframe
s2_seehear <- s2_allq_allf %>%
  select(IndivID, pq3, pq3b, pq3c,
         pq6, pq6b, pq6c, pq7, pq7a, pq7b,
         tellegentotal, condition, playcount) %>% 
  #rename(mentalimage = pq3, stronger = pq3b, eyes = pq3c) %>%
  #pivot_longer(cols = c(mentalimage, stronger, eyes), names_to = "question", values_to = "response")
  mutate(see_total = rowSums(across(pq3:pq3c), na.rm = T), 
         hear_total = rowSums(across(pq6:pq7), na.rm=T)) %>% 
  pivot_longer(cols = c(see_total, hear_total), names_to = "modality", values_to = "total")


# t-tests

##difference in seeing?
s2_seehear %>% filter(modality == "see_total") %>%
    group_by(condition) %>%
    get_summary_stats(total, type = "mean_sd")

t_test(total ~ condition, 
       data = s2_seehear %>% filter(modality == "see_total"))

##difference in hearing?
s2_seehear %>% filter(modality == "hear_total") %>%
    group_by(condition) %>%
    get_summary_stats(total, type = "mean_sd")

t_test(total ~ condition, 
       data = s2_seehear %>% filter(modality == "hear_total"))
```

`r as.numeric(count(s2_seehear %>% filter(condition == "seeing", modality == "see_total", total>0)))` participants in the seeing condition answered yes to at least one seeing specific question, whereas only `r as.numeric(count(s2_seehear %>% filter(condition == "hearing", modality == "see_total", total>0)))` did. 

Conversely, `r as.numeric(count(s2_seehear %>% filter(condition == "hearing", modality == "hear_total", total>0)))` people in the hearing condition answered yes to at least on hearing specific question, whereas only `r as.numeric(count(s2_seehear %>% filter(condition == "seeing", modality == "hear_total", total>0)))` people in the seeing condition did. 

## Correlation of Leland Scores and Absorption

```{r}
# correlation between absorption and leland scores (conditions combined)
pearson_r <- cor.test(s2$tellegentotal, s2$leland_score,
                             method = "pearson")
pearson_r
```

## Model 

### Model as reported in Study 2 Results 

```{r}
# model where score is predicted by group and absorption (and participant?)

## data frame (all items, scaled)
s2_m <- s2 %>% 
  select(IndivID, condition, leland_score, playcount, tellegentotal) %>% 
  mutate(IndivID = factor(IndivID))

## linear model
s2.m <- lm(leland_score ~ scale(tellegentotal, scale=F) + playcount + condition, 
           data = s2_m)

parameters(s2.m)
```

### Model results using expanded Leland scores (as reported in Appendix B)

```{r}
# models with expanded scores (as reported in Appendix B)

## data frame (all items, scaled)
s2_m_allq <- s2_allq %>% 
  select(IndivID, condition, leland_score, playcount, tellegentotal) %>% 
  mutate(IndivID = factor(IndivID))

## linear model
s2.m_allq <- lm(leland_score ~ scale(tellegentotal, scale=F) + playcount + condition, 
           data = s2_m_allq)

parameters(s2.m_allq)

#---

## data frame (all items, scaled)
s2_m_allq_allf <- s2_allq_allf %>% 
  select(IndivID, condition, leland_score_plus, playcount, tellegentotal) %>% 
  mutate(IndivID = factor(IndivID))

## linear model
s2.m_allq_allf <- lm(leland_score_plus ~ scale(tellegentotal, scale=F) + playcount + condition, 
           data = s2_m_allq_allf)

parameters(s2.m_allq_allf)

```

## Plots

```{r}
#scores plotted against absorption  

## all together
s2 %>% 
  select(IndivID, leland_score, condition, tellegentotal) %>% 
  ggplot(aes(x=tellegentotal, y=leland_score, color = condition)) + 
  geom_jitter(
    aes(color = condition),
    height = 0, width = 0, 
              size = 1, alpha = 1) +
  geom_smooth(aes(condition = NULL), color = "black", method = "lm") +
  scale_color_brewer(palette = "Set2") +
  labs(title = "Relationship between Leland Score and Absorption Score",
       subtitle = "Both Scores Summed",
       x = "Absorption Score",
       y = "Leland Score",
       color = "Condition") +
   guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)))

## faceted
s2 %>% 
  select(IndivID, leland_score, condition, tellegentotal) %>% 
  ggplot(aes(x=tellegentotal, y=leland_score)) + 
  geom_jitter(
    aes(color = condition),
    height = 0, width = 0, 
              size = 1, alpha = 1) +
  geom_smooth(aes(condition = NULL), color = "black", method = "lm") +
  facet_wrap(~ condition, ncol = 2) + 
  scale_color_brewer(palette = "Set2") +
  labs(title = "Relationship between Leland Score and Absorption Score",
       subtitle = "Both Scores Summed",
       x = "Absorption Score",
       y = "Leland Score",
       color = "Condition") +
   guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)))

```

## Student comments

```{r}
# These are student responses to the question "Did you have any other unusual experience that you associated with Leland?" 
## Quotes from these responses were selected and reported in the manuscript 

s2_clean %>% select(comments) %>% rename(`Did you have any other unusual experience that you associated with Leland?` = comments) %>% unique()
```
